---
title: "exo12"
format: html
editor: visual
---

## Step 1

Load data and change the value of-999

```{r}
library(tidyverse)
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f, col_names = TRUE)
head(d)

install.packages("naniar")
library(naniar)

d <- replace_with_na_all(d, condition = ~.x == -999)

```

## Step 2

Drop the variable litter size and refs

```{r}

d <- d[, !(names(d) %in% c("litter size", "refs"))]
head(d)
```

## Step 3

Log all of the variables

```{r}

library(dplyr)
d <- d %>% 
  mutate(
  "logMass" = log(`mass(g)`),
  "logGestation" = log(`gestation(mo)`),
  "logNewBorn" = log(`newborn(g)`),
  "logWeaning" = log(`weaning(mo)`),
  "logWeanMass" = log(`wean mass(g)`),
  "logAFR" = log(`AFR(mo)`),
  "logMaxLife" = log(`max. life(mo)`),
  "logLitterYear" = log(`litters/year`)
)
```

## Step 4

Regression and add residuals

```{r}
#Regression
model <- lm(logMass ~ logGestation + logWeaning + logAFR + logMaxLife + logNewBorn + logWeanMass, data = d)


# Extract residuals and add them to the dataframe
library(dplyr)
d <- mutate(d,
            relGest = residuals(lm(logMass ~ logGestation, data = d, na.action = na.exclude)),
            relWean = residuals(lm(logMass ~ logWeaning, data = d, na.action = na.exclude)),
            relAFR = residuals(lm(logMass ~ logAFR, data = d, na.action = na.exclude)),
            relLife = residuals(lm(logMass ~ logMaxLife, data = d, na.action = na.exclude)),
            relNewbornMass = residuals(lm(logMass ~ logNewBorn, data = d, na.action = na.exclude)),
            relWeaningMass = residuals(lm(logMass ~ logWeanMass, data = d, na.action = na.exclude))
)
head(d)


```

## Step 5

Plots

```{r}
#Step 5
#plot

#max lifespan in relation to order

library(ggplot2)

ggplot(d, aes(x = order, y = relLife)) +
  geom_boxplot() +
  labs(x = "Mammalian Order", y = "Residuals of Max Lifespan") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  

#newborn mass in relation to order

ggplot(d, aes(x = order, y = relNewbornMass)) +
  geom_boxplot() +
  labs(x = "Mammalian Order", y = "Residuals of Newborn Mass") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#weaning mass in relation to order

ggplot(d, aes(x = order, y = relWeaningMass)) +
  geom_boxplot() +
  labs(x = "Mammalian Order", y = "Residuals of Weaning Mass") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Step 6

Plots

```{r}
install.packages("tidymodels")
library(tidymodels)

# Drop NA
d_new <- d %>%
  drop_na(logMaxLife, logGestation, logNewBorn, logWeaning, logWeanMass, logLitterYear, logMass, logAFR)

# Predictor: max life (I choose to drop variables as I go)
library(MASS)
m0 <- lm(logMaxLife ~ logGestation + logNewBorn + logWeaning + logWeanMass + logLitterYear + logMass, data = d_new)
summary(m0)
m1 <- update(m0, . ~ . - logNewBorn)
drop1(m1, test = "F")

m2 <- update(m1, . ~ . - logWeanMass)
drop1(m2, test = "F")

m3 <- update(m2, . ~ . - logWeaning)
drop1(m3, test = "F")

m4 <- update(m3, . ~ . - logMass)
drop1(m4, test = "F")

m5 <- update(m4, . ~ . - logLitterYear)
drop1(m5, test = "F")

s1 <- stepAIC(m0, scope = . ~ ., direction = "both")
summary(s1)

library(AICcmodavg)
aictab(list(m0, m1, m2, m3, m4, m5), c("m0", "m1", "m2", "m3", "m4", "m5"))

# Predictor: AFR
ma <- lm(logAFR ~ logGestation + logNewBorn + logWeaning + logWeanMass + logLitterYear + logMass, data = d_new)
summary(ma)

mb <- update(ma, . ~ . - logNewBorn)
drop1(mb, test = "F")

mc <- update(mb, . ~ . - logWeanMass)
drop1(mc, test = "F")

md <- update(mc, . ~ . - logMass)
drop1(md, test = "F")

me <- update(md, . ~ . - logWeaning)
drop1(me, test = "F")

mf <- update(me, . ~ . - logLitterYear)
drop1(mf, test = "F")

s2 <- stepAIC(ma, scope = . ~ ., direction = "both")
summary(s2)

```

Top variables

For Maximum lifespan, m2, m3, m0 are the best models and for age at first reproduction mc, md, mb and me are the best models

```{r}
library(AICcmodavg)
aictab(list(ma, mb, mc, md, me, mf), c("ma", "mb", "mc", "md", "me", "mf"))#I just list the models with their Delta_AICc

# Top models
top1 <- list(m2, m1, m0)
top2 <- list(mc, md, mb, me)
top1
top2

```

Gestation appear in all top model and weaning followed by litters size per year are other variables that mostly appear in the top models.

Calculate and plot the model-averaged coefficient

```{r}
#  model-averaged coefficients for top1
library(MuMIn)

modavg_coefs1 <- coef(model.avg(top1), se = TRUE)
modavg_coefs2 <- coef(model.avg(top2), se = TRUE)

#plot
str(modavg_coefs1)
plot(
  modavg_coefs1,
  ylim = c(-2, 2),  
  pch = 16,
  xlab = "Coefficient",
  ylab = "Model Average",
  main = "Model-Averaged Coefficients (Top 1)"
)
plot(
  modavg_coefs2,
  ylim = c(-2, 2),  # Adjust ylim as needed
  pch = 16,
  xlab = "Coefficient",
  ylab = "Model Average",
  main = "Model-Averaged Coefficients (Top 2)"
)
```
